Regional Base-Flow Index in Arid Landscapes Using Machine Learning and Instrumented Records

### 2025-05-01 elsevier_MANUSCRIPT.qmd
Update elsevier_MANUSCRIPT.qmd

Daily observed streamflow data obtained from the United States Geological Survey (USGS) National Water Information System (NWIS) were used in this study. Streamgages were selected depending on criteria to ensure the applicability of each site. Following the findings of @odonnell2016, which determined that 8–10 years of calibration data are necessary to account for climate variability in paired watershed studies in the region, a minimum record length of 10 years was required. Additionally, years with more than 30 missing days of streamflow data were excluded from the analysis. Streamgages directly downstream of major regulation (e.g., reservoirs, lakes, diversions) were excluded, based on USGS annual water data reports and site metadata \[@usgs2010\]. While some flow alteration is widespread, the focus was on removing gauges with clear, immediate regulatory impacts. As such, streamgages along the Colorado River were omitted because they represent managed flows governed by the Colorado River Compact. After applying these selection criteria, 205 USGS streamgages with acceptable periods of record were included in the study (@fig-study-area). Periods of record ranged from 10 to 112 years, with a median of 28 years.

@@

To check for autocorrelation, we applied the Durbin-Watson test [@durbin1950testing], which revealed significant autocorrelation at four streamgages on an annual basis. Of these, only one streamgage (09486500 - Santa Cruz River at Cortaro, AZ) showed a significant trend in BFI. This streamgage was excluded from the trend analysis, as autocorrelation could inflate the variance of the Mann-Kendall statistic, potentially leading to biased trend estimates [@MK_Hamed1998]. Trends with a $\rho \le 0.05$ are considered significant.

>>>>

Daily observed streamflow data obtained from the United States Geological Survey (USGS) National Water Information System (NWIS) were used in this study. Streamgages were selected depending on criteria to ensure the applicability of each site. Following the findings of @odonnell2016, which determined that 8–10 years of calibration data are necessary to account for climate variability in paired watershed studies in the region, a minimum record length of 10 years was required. Additionally, years with more than 30 missing days of streamflow data were excluded from the analysis. Streamgages directly downstream of major regulation (e.g., reservoirs, lakes, diversions) were excluded, based on USGS annual water data reports and site metadata [@usgs2010]. While some flow alteration is widespread, the focus was on removing gauges with clear, immediate regulatory impacts. As such, streamgages along the Colorado River were omitted because they represent managed flows governed by the Colorado River Compact. After applying these selection criteria, 205 USGS streamgages with acceptable periods of record were included in the study (@fig-study-area). Periods of record ranged from 10 to 112 years, with a median of 28 years.

@@

To check for autocorrelation, we applied the Durbin-Watson test [@durbin1950testing], which revealed significant autocorrelation at four streamgages on an annual basis. Of these, only one streamgage (09486500 - Santa Cruz River at Cortaro, AZ) showed a significant trend in BFI. This streamgage was excluded from the trend analysis, as autocorrelation could inflate the variance of the Mann-Kendall statistic, potentially leading to biased trend estimates [@MK_Hamed1998]. Trends with a $\alpha \le 0.05$ are considered significant.
